---
title: "liuWestV2"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This R Markdown allows us to play around a little with Nimble based on https://r-nimble.org/html_manual/cha-algos-provided.html

First we will build our model. This is our state space model... which for now has random parameters.

```{r}
#Load nimble packages
library(nimble)
library(nimbleSMC)

# Building a simple linear state-space model. 
# x is latent space, y is observed data
timeModelCode <- nimbleCode({
  x[1] ~ dnorm(mu_0, 1)
  y[1] ~ dnorm(x[1], 1)
  for(i in 2:t){
    x[i] ~ dnorm(x[i-1]*a + (1-x[i-1])*b, 0.01)
    y[i] ~ dnorm(x[i]*c*cos(3.14*i)+(1-x[i])*d*cos(i), 1)
  }
  
  a ~ dunif(0, 1)
  b ~ dunif(0, 1)
  c ~ dnorm(1,1)
  d ~ dnorm(1,1)
  mu_0 ~ dnorm(0, 1)
})
```

Now we will generate some fake data using the above state space model.

```{r}
set.seed(2)
t <- 100; mu_0 <- 1
x <- rbinom(1,1, mu_0)
y <- rnorm(1, x, 1)
a <- 0.5; b <- 1; c <- 1
for(i in 2:t){
  x[i] <- rbinom(1,1,x[i-1]* a + (1-x[i-1])*b )# rnorm(1, x[i-1] * a + b, 1)
  y[i] <- rnorm(1, x[i]*1*c+2*(1-x[i]), 1)
}
```

Under normal conditions, we could compile it and then use, e.g., MCMC on the compiled model. Note that, in the below, we are setting the unknown parameters in the state space models to known scalars. 

```{r}

# build the model
rTimeModel <- nimbleModel(timeModelCode, constants = list(t = t), 
                          data <- list(y = y), check = FALSE )

cTimeModel <- compileNimble(rTimeModel)
```

However, we aren't interested in MCMC here... we want to use a filtering algorithm. 
## Liu West

We will now do the Liu West filter which will allow us to do inference on the unknown (but constant) parameters in our transition and observation models.
```{r}
num_particles = 100
LWTimeModel <- rTimeModel$newModel(replicate = TRUE)
compileNimble(LWTimeModel)
```



```{r}
# Build Liu-West filter, also 
# specifying which top level parameters to estimate
rLWF <- buildLiuWestFilter(LWTimeModel, "x", params = c("a", "b", "c", "d"),
                           control = list(saveAll = TRUE)) 
```

```{r}
#rLWF$run(10)
cLWF <- compileNimble(rLWF,project = LWTimeModel)
```


```{r}
# Run Liu-West filter
cLWF$run(num_particles)
rLWF$run(num_particles)
```


```{r}
#cLWF$run(100)
#rLWF$run(100)
#matrix(unlist(rLWF$mvEWSamples$x), nrow=num_particles, byrow=T)
particles_x = matrix(unlist(rLWF$mvEWSamples$x),nrow=num_particles,byrow=T)
particles_a = matrix(unlist(rLWF$mvEWSamples$a),nrow=num_particles,byrow=T)
particles_b = matrix(unlist(rLWF$mvEWSamples$b),nrow=num_particles,byrow=T)
particles_c = matrix(unlist(rLWF$mvEWSamples$c),nrow=num_particles,byrow=T)
particles_d = matrix(unlist(rLWF$mvEWSamples$d),nrow=num_particles,byrow=T)
hist(particles_d)
```

##Check Filtering Dist of Latent Variables
```{r}
mean_vec = c()
for (i in 1:t){
  mean_vec[i]= mean(particles_x[,i])
}
plot(1:t,mean_vec,type='l', col='blue')
lines(1:t, x)
```


