---
title: "Nimble Playground"
output: html_document
date: "2025-04-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This R Markdown allows us to play around a little with Nimble based on https://r-nimble.org/html_manual/cha-algos-provided.html

First we will build our model. This is our state space model... which for now has random parameters.

```{r}
#Load nimble packages
library(nimble)
library(nimbleSMC)

# Building a simple linear state-space model. 
# x is latent space, y is observed data
timeModelCode <- nimbleCode({
  x[1] ~ dnorm(mu_0, 1)
  y[1] ~ dnorm(x[1], 1)
  for(i in 2:t){
    x[i] ~ dnorm(x[i-1] * a + b, 1)
    y[i] ~ dnorm(x[i] * c, 1)
  }
  
  a ~ dunif(0, 1)
  b ~ dnorm(0, 1)
  c ~ dnorm(1,1)
  mu_0 ~ dnorm(0, 1)
})
```

Now we will generate some fake data using the above state space model.

```{r}

# simulate some data
t <- 25; mu_0 <- 1
x <- rnorm(1 ,mu_0, 1)
y <- rnorm(1, x, 1)
a <- 0.5; b <- 1; c <- 1
for(i in 2:t){
  x[i] <- rnorm(1, x[i-1]*  a + b, 1)
  y[i] <- rnorm(1, x[i] * c, 1)
}
```

Under normal conditions, we could compile it and then use, e.g., MCMC on the compiled model. Note that, in the below, we are setting the unknown parameters in the state space models to known scalars. 

```{r}

# build the model
rTimeModel <- nimbleModel(timeModelCode, constants = list(t = t), 
                          data <- list(y = y), check = FALSE )

# Set parameter values and compile the model
rTimeModel$a <- 0.5
rTimeModel$b <- 1
rTimeModel$c <- 1
rTimeModel$mu_0 <- 1

cTimeModel <- compileNimble(rTimeModel)
```

However, we aren't interested in MCMC here... we want to use a filtering algorithm. 

## Bootstrap Filter

We start with the bootstrap filter. We build and compile the model first. 

```{r}

# Build bootstrap filter
rBootF <- buildBootstrapFilter(rTimeModel, "x", 
                               control = list(thresh = 0.8, saveAll = TRUE, 
                                              smoothing = FALSE))
# Compile filter   
cBootF <- compileNimble(rBootF,project = rTimeModel)
```

Thereafter, we can run it to get some information on it. 

```{r}
#Get Filtering Dist
num_particles = 100
rBootF$run(num_particles) #Run 5 particles thru the model
particle_paths = rBootF$mvEWSamples$x  #Gives equally weighted samples from the posterior
#mvWSamples gives weighted samples
filtering_sample = matrix(0,nrow=num_particles,ncol=t)
mean_filtering_dist = c()

#Build data frame showing particle's path throughout the time steps... can be used to make filtering dist
for(j in 1:t){
  for(i in 1:num_particles){
    filtering_sample[i,j] = particle_paths[[i]][j]}
 mean_filtering_dist[j] = mean(filtering_sample[,j])
}
hist(filtering_sample[,t])
print(c('True x(25) = ', x[t] ))
```

We can also get some summary statistics from the model that may make things easier when fitting models. 

```{r}
# Run bootstrap filter, which returns estimate of model log-likelihood
bootLLEst <- rBootF$lastLogLik
bootESS <- rBootF$returnESS()
```

To get an idea of how well it's working, as a rough hack, we can track the mean of our filtering distribution versus truth. 

```{r}
lb = min(min(mean_filtering_dist), min(x)) -1
ub = max(max(mean_filtering_dist), max(x)) +1
plot(1:t, mean_filtering_dist, type='l', ylim = c(lb,ub), col='blue')
lines(1:t,x)
```

## Liu West

We will now do the Liu West filter which will allow us to do inference on the unknown (but constant) parameters in our transition and observation models. 

```{r}
num_particles = 100
LWTimeModel <- rTimeModel$newModel(replicate = TRUE)
compileNimble(LWTimeModel)
# Build Liu-West filter, also 
# specifying which top level parameters to estimate
rLWF <- buildLiuWestFilter(LWTimeModel, "x", params = c("a", "b", "c"),
                           control = list(saveAll = TRUE))   
#rLWF$run(10)
cLWF <- compileNimble(rLWF,project = LWTimeModel)
# Run Liu-West filter
cLWF$run(num_particles)
rLWF$run(num_particles)
```


```{r}
#cLWF$run(100)
#rLWF$run(100)
#matrix(unlist(rLWF$mvEWSamples$x), nrow=num_particles, byrow=T)
particles_x = matrix(unlist(rLWF$mvEWSamples$x),nrow=num_particles,byrow=T)
particles_a = matrix(unlist(rLWF$mvEWSamples$a),nrow=num_particles,byrow=T)
particles_b = matrix(unlist(rLWF$mvEWSamples$b),nrow=num_particles,byrow=T)
particles_c = matrix(unlist(rLWF$mvEWSamples$c),nrow=num_particles,byrow=T)
hist(particles_c)
```

##Check Filtering Dist of Latent Variables
```{r}
mean_vec = c()
for (i in 1:t){
  mean_vec[i]= mean(particles_x[,i])
}
plot(1:t,mean_vec,type='l', col='blue')
lines(1:t, x)
```


